---
title: "Simulate RNAseq counts"
output: html_document
date: "2024-05-07"
---

```{r}
setwd("/Users/katsiarynadavydzenka/Documents/PhD_AI/CN-aware-DGE/simulation")

library(tidyverse)
library(compcodeR)
library(DESeq2)
```

## Using compcodeR simulator


```{r}
rna_counts_sim <- compcodeR::generateSyntheticData(dataset = "rna_counts_sim", n.vars = 12000, 
                                                   samples.per.cond = 20, 
                                                   n.diffexp = 1400, 
                                                   repl.id = 1, seqdepth = 1e7, 
                                                   fraction.upregulated = 700, 
                                                   #between.group.diffdisp = FALSE, 
                                                   filter.threshold.total = 1, 
                                                   filter.threshold.mediancpm = 0, 
                                                   fraction.non.overdispersed = 0, 
                                                   relmeans = mean,
                                                   dispersions = dispersion,
                                                   random.outlier.high.prob = 0,
                                                   random.outlier.low.prob = 0,
                                                   effect.size = 1.5,
                                                   output.file = "rna_counts_sim.rds")

rna_counts_sim <- rna_counts_sim@count.matrix
#rna_normal <- luad_rna_sd %>% as.data.frame() %>% select(1:45)
#rna_tumor <- luad_rna_sd %>% as.data.frame() %>% select(46:90)
#rna_counts <- cbind(brca_rna_normal, brca_rna_tum)
```

## Using DESeq2 simulator function

```{r}
library(DESeq2)

deseq_sim <- makeExampleDESeqDataSet(
  n = 12000,
  m = 40,
  betaSD = 0,
  interceptMean = 6,
  interceptSD = 2,
  dispMeanRel = function(x) 6/x + 0.6
  #sizeFactors = rep(1, m)
)
```
item{n}{number of rows}
item{m}{number of columns}
item{betaSD}{the standard deviation for non-intercept betas, i.e. beta ~ N(0,betaSD)}
item{interceptMean}{the mean of the intercept betas (log2 scale)}
item{interceptSD}{the standard deviation of the intercept betas (log2 scale)}
item{dispMeanRel}{a function specifying the relationship of the dispersions on code{2^trueIntercept}}
item{sizeFactors}{multiplicative factors for each sample}

```{r}
# Generate metadata
metadata <- data.frame(patID = colnames(rna_counts),
                       condition = rep(c("0", "1"), each = 20))
metadata <- metadata %>% remove_rownames %>% column_to_rownames(var = "patID") 
```

```{r}
# DESeq2 fit
dds <- DESeqDataSetFromMatrix(countData = rna_cnv,
                              colData = metadata,
                              design= ~ condition)
dds <- DESeq(dds)
```
```{r}
#extract results from dds object
res <- results(dds1, name="condition_B_vs_A") %>% as.data.frame()
```
```{r}
rna_dds_sim <- deseq_sim@assays@data@listData[["counts"]]
rownames(rna_dds_sim) <- paste0("G", 1:(nrow(rna_dds_sim)))
rna_normal <- rna_dds_sim[,1:20] %>% as.data.frame()
rna_tumor <- rna_dds_sim[,21:40] %>% as.data.frame()
rna_counts <- cbind(rna_normal, rna_tumor)
```


```{r}
# CNV simulation
cnv_tumor <- as.matrix(luad_cnv_sd)
cnv_tumor_sd <- matrixStats::colSds(cnv_tumor) %>% 
  as.data.frame() %>% 
  setNames("sd") %>% 
  subset(sd > 1.8)

cnv_tumor <- cnv_tumor[,colnames(cnv_tumor) %in% rownames(cnv_tumor_sd)]
cnv_tumor <- cnv_tumor[1:3000,]
cnv_tumor <- replace(cnv_tumor, cnv_tumor>10, 10)

colnames(cnv_tumor) <- colnames(rna_tumor)
rownames(cnv_tumor) <- paste0("G", 1:(nrow(cnv_tumor)))

cnv_0 <- sapply(1:20, function(x) sample(x=c(0.5,1,2,3), size = 300, replace=TRUE, prob = c(.50, .30, .10, .10)))
cnv_1 <- sapply(1:20, function(x) sample(x=c(0.5,1,2,3), size = 300, replace=TRUE, prob = c(.20, .60, .10, .10)))
cnv_2 <- sapply(1:20, function(x) sample(x=c(1,2,3,4), size = 8400, replace=TRUE, prob = c(.05, .70, .10, .10)))
cnv_tumor <- rbind(cnv_tumor, cnv_0, cnv_1, cnv_2) %>% as.matrix()
cnv_norm <- matrix(2, nrow(rna_normal), 20)

colnames(cnv_norm) <- colnames(rna_normal)
rownames(cnv_norm) <- rownames(rna_normal)

cnv <- cbind(cnv_norm, cnv_tumor)
cnv <- apply(cnv, 2, function(x) x/2)

rownames(cnv) <- paste0("G", 1:(nrow(cnv)))
```


```{r}
cnv <- cbind(cnv_tumor, cnv_norm)
cnv <- apply(cnv, 2, function(x) x/2)
rna_cnv <- rna_mixed * cnv
rna_cnv <- ceiling(rna_cnv)
```
```{r}
dds <- estimateSizeFactors(deseq_sim)
plotSparsity(dds)

```

## Using OmicsSIMLA

```{r}
rna_counts_sim1 <- read.table("rna_luad1.tsv", header = TRUE, sep = "")
rna_counts_sim2 <- read.table("rna_luad2.tsv", header = TRUE, sep = "")

rna_counts_sim2 <- rna_counts_sim2[,c(1:5000)]
rna_counts_sim2 <- rna_counts_sim2 %>% remove_rownames %>% column_to_rownames(var="FAM")
colnames(rna_counts_sim) <- paste0("S", 1:(ncol(rna_counts_sim)))
rownames(rna_counts_sim) <- paste("G", 1:(nrow(rna_counts_sim)))
```

```{r}
## Prepare CN data ##
cnv <- apply(cnv, 1, function(x) x/2)
cnv <- apply(cnv, 1, function(x) x+10e-9)

rna_cnv <- rna_nocnv * cnv
```
```{r}
# Simulation of CN independent expression
rna_tumor <- rna_counts[,21:40]
rna_normal <- rna_counts[,1:20]
#rna_up <- rna_tumor[1:300,]*2.5
rna_cn <- rna_tumor[1:3600,]
#rna_down <- rna_tumor[3301:3600,]/2.5
rna_oth_up <- rna_tumor[3601:4600,]*2
rna_oth_down <- rna_tumor[4601:5600,]/2
rna_dip <- rna_tumor[5601:12000,]
rna_tumor_mix <- rbind(rna_cn, rna_oth_up, rna_oth_down, rna_dip)
rna_mixed <- cbind(rna_normal, rna_tumor_mix)
rna_mixed <- ceiling(rna_mixed)

```

